---
title: "Exploratory Data Analysis"
author: "Adam Whalen"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)


knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 0.6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.color = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_color_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Load the weather data

```{r load}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())

weather_df
```

## `group_by`

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  ungroup(month) # <-- you can also ungroup things
```

## Counting things

Let's count month / name observations.

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(n_obs = n())
```

We can use `count` as well.

```{r}
weather_df %>% 
  count(month, name = "n_obs")
```

**NEVER** use base R's `table`.

```{r baser_table, eval = FALSE}
weather_df %>% 
  pull(month) %>% 
  table()
```

Other helpful counters:

```{r}
weather_df %>% 
  group_by(month) %>% 
  summarize(
    n_obs = n(),
    n_days = n_distinct(date)
  )
```

## A digression on 2x2 tables

```{r}
weather_df %>% 
  filter(name != "Waikiki_HA") %>% 
  mutate(
    cold = case_when(
      tmax < 5  ~ "cold",
      tmax >= 5 ~ "not cold",
      TRUE      ~ ""
    )
  ) %>% 
  group_by(name, cold) %>% 
  summarize(count = n())
```

We have the data, but let's get it into a 2x2 table looking thing.

```{r}
weather_df %>% 
  filter(name != "Waikiki_HA") %>% 
  mutate(
    cold = case_when(
      tmax < 5  ~ "cold",
      tmax >= 5 ~ "not cold",
      TRUE      ~ ""
    )
  ) %>% 
  janitor::tabyl(name, cold)
```

NB: May not actually need to build a 2x2 table very often (likely more useful to get an OR via regression). But if you really need one, use `janitor::tabyl`.

## General summaries

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE),
    mean_prcp = mean(prcp, na.rm = TRUE),
    median_tmin = median(tmin, na.rm= TRUE)
  ) 
```

This is a dataframe! Which means, we can make a plot!

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE),
    mean_prcp = mean(prcp, na.rm = TRUE),
    median_tmin = median(tmin, na.rm= TRUE)
  ) %>% 
  ggplot(aes(x = month, y = mean_tmax, color = name)) +
  geom_point() +
  geom_line()
```

Suppose you want to summarize many columns.

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(across(prcp:tmin, mean))
```

Reminder: sometimes your results are easier to read in another format.

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(mean_tmax = mean(tmax)) %>% 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax
  ) %>% 
  knitr::kable(digits = 1)
```

## `group_by` and `mutate`

```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(
    mean_tmax = mean(tmax, na.rm = TRUE),
    centered_tmax = tmax - mean_tmax
  )
```

`mutate` adds my group/summary variable to the existing data frame, rather than creating a new one. Helpful when you want to center/standardize values!

Plot for illustration:

```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(
    mean_tmax = mean(tmax, na.rm = TRUE),
    centered_tmax = tmax - mean_tmax
  ) %>% 
  ggplot(aes(x = date, y = centered_tmax, color = name))+
  geom_point()
```

What about window functions?

Ranking...

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  mutate(temp_rank = min_rank(tmax))
```

Lag: shift values down by one (or a specified number of) observation(s)

```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(lag_temp = lag(tmax))
```

```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(temp_change = tmax - lag(tmax)) %>% 
  summarize(
    temp_change_max = max(temp_change, na.rm = TRUE),
    temp_change_sd = sd(temp_change, na.rm = TRUE)
  )
```

## Quick note

`summarize` only gets you so far. Can't really do more than single numbers (univariate analysis). 


