---
title: "Visualization"
subtitle: "Part 2"
author: "Adam Whalen"
output: github_document
---

```{r setup}
library(tidyverse)
library(patchwork)
```

## Load the weather data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df
```

## Remember this plot?

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5)
```

## Labels
It's helpful to add in captions to describe the plot.

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs (
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017"
  )
```

## Scales

Let's start with the same plot; x and y scales

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs (
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017"
  ) +
  scale_x_continuous(
    breaks = c(-15, 0, 15),
    labels = c("-15 C", "0 C", "15 C")
  ) +
  scale_y_continuous(
    # trans = "log" 
    # position = "right"
  )
```

Now, let's try color scales. FIrst making our own (bad)

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs (
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017"
  ) +
  scale_color_hue(
    name = "Location", # rename the legend
    h = c(100, 300))
```

...then using a built color scale.

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs (
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017"
  ) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE
    )
```

## Themes

Shift the legend.

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017"
  ) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE
    ) +
  theme(legend.position = "bottom")
```

Now, change the overall theme.

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017"
    ) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE
    ) +
  # theme_bw()
  # theme_classic()
  theme_minimal() +
  theme(legend.position = "bottom")
```

Look around in `ggthemes::` for different themes, like `theme_economist` or `theme_excel`. As long as it's clear, it doesn't really matter. Note that theme tweaks (like legend position) have to be set after the main theme is set, otherwise it is overwritten.

## Setting options

It can be helpful to include global options at the beginning of every working document that will be consistent throughout.

```{r, eval = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 0.6
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.color = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_color_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Data args in `geom`

```{r}
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_line(data = central_park)
```

It can be helpful to know that you can splice and then reshare two datasets within the same graph (inside different geoms).

## `patchwork`

Remember faceting?

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = 0.5) +
  facet_grid(. ~ name)
```

Faceting allows us to separate panes within the same geometry and aesthetic mappings. But what happens when you want multipanel plots but can't facet?

```{r}
tmax_tmin_p = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  theme(legend.position = "none")

prcp_dens_p = 
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = 0.5) +
  theme(legend.position = "none")

tmax_date_p = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none")

(tmax_tmin_p + prcp_dens_p) / tmax_date_p
```

## Data manipulation

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_violin(alpha = 0.5)
```

Why do the violins get arranged like that? Because the names are stored as characters, but `ggplot` converts them to vectors, assigning them numbers by alphabetical order. Let's change that.

```{r}
weather_df %>% 
  mutate(
    name = as.factor(name),
    name = forcats::fct_relevel(name, c("Waikiki_HA"))
    ) %>% 
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_violin(alpha = 0.5)
```

What if I wanted densities for `tmin` and `tmax` simultaneously?

```{r}
weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  pivot_longer(
    tmax:tmin,
    names_to = "observation",
    values_to = "temps"
  ) %>% 
  ggplot(aes(x = temps, fill = observation))+
  geom_density(alpha = 0.5)
```

Or, instead of filtering, facet by name.

```{r}
weather_df %>% 
  pivot_longer(
    tmax:tmin,
    names_to = "observation",
    values_to = "temps"
  ) %>% 
  ggplot(aes(x = temps, fill = observation))+
  geom_density(alpha = 0.5) +
  facet_grid(. ~ name)
```

## Revisit the pups

Now, for a bonus (hard) plotting example. Back to the FAS study.

```{r}
pup_data = 
  read_csv("./data/FAS_pups.csv") %>% 
  janitor::clean_names() %>% 
  mutate(sex = recode(sex, `1` = "male", `2` = "female"))

litters_data = 
  read_csv("./data/FAS_litters.csv") %>% 
  janitor::clean_names() %>% 
  separate(group, into = c("dose", "day_of_tx"), sep = 3)

fas_data = left_join(pup_data, litters_data, by = "litter_number")

fas_data %>% 
  ggplot(aes(x = dose, y = pd_ears)) +
  geom_violin() +
  facet_grid(. ~ day_of_tx)
```

Being able to show all four outcomes, in a meaningful and clear order, is less of a `ggplot` problem and more of a data manipulation problem.

```{r}
fas_data %>% 
  select(dose, day_of_tx, starts_with("pd_")) %>% 
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome",
    values_to = "pn_day"
  ) %>% 
  drop_na() %>% 
  mutate(outcome = forcats::fct_relevel(outcome, "pd_ears", "pd_pivot", "pd_walk", "pd_eyes")) %>% 
  ggplot(aes(x = dose, y = pn_day)) +
  geom_violin() +
  facet_grid(day_of_tx ~ outcome)
```







